---
title: "ms6360 - Homework 1"
author: "Maisie Sun"
date: 2022-09-24
output: github_document
---
# Problem 1

```{r, message = FALSE}
library(tidyverse)
data("penguins", package = "palmerpenguins")
```

Here are some descriptions of the `penguins` dataset:

* There are 344 observations of 8 variables (`r nrow(penguins)` rows and `r ncol(penguins)` columns).

* Each row depicts a penguin subject and there are 344 penguins studied.

* The 8 variables that were measured are: species, island, bill length in mm, bill depth in mm, flipper length in mm, body mass in g, sex, and year of data collection.

* The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm=TRUE), 2)`.


Now, onto the scatterplot
```{r, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave("penguins_scatter_plot.pdf", heigh = 4, width = 6)
```

# Problem 2

Creating a dataframe to include:

* a random sample of size 10 from a standard Normal distribution

* a logical vector indicating whether elements of the sample are greater than 0

* a character vector of length 10

* a factor vector of length 10, with 3 different factor “levels”

```{r}
my_df = 
  tibble(
    random = rnorm(n = 10),
    vec_logical = random > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vec_factor = factor(c("high", "middle", "low", "low", "high", "middle", "high", "low", "middle", "high"))
    )
my_df

mean(pull(my_df, random))
mean(pull(my_df, vec_logical))
mean(pull(my_df, vec_char)) 
mean(pull(my_df, vec_factor))

```

Running the code, numeric and logical vectors work and the means were calculated. Character and factor vectors did not work and the mean was not calculated. For the warning message for character and factor vectors is "argument is not numeric or logical: returning NA".

When explicitly converting a logical, character, and factor variable to numeric, it looks like this:

```{r, results = "hide", warning = FALSE}
as.numeric(pull(my_df, vec_logical))
as.numeric(pull(my_df, vec_char))
as.numeric(pull(my_df, vec_factor))
```

After turning the logical, character, and factor variable into numeric, only the character variable was unable to be turned into numeric. The warning message shows: "NAs introduced by coercion".

This explains why the character variable cannot have the mean calculated because it cannot be converted into a numeric variable.

The factor variable can now have its mean being calculated, being `mean(as.numeric(pull(my_df, vec_factor)))`, in contrast to the first mean calculation which was impossible. This shows that R is unable to recognize an ordinal variable once it is converted into a numeric variable. This is something to be careful when dealing with ordinal variables.